---
title: "California Cetaceans Species Richness"
description: |
  Spatial analysis of species richness for 35 cetacean species along the California Bight.
author:
  - name: Anastasia Kunz
date: 02-23-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
    theme: flatly
    code_folding: hide
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(raster)
library(tidyverse)
library(here)
library(sf)
library(fasterize)
library(rnaturalearth)
library(sp)

```


## Cetacean Species Along the California Bight

Species richness is a measure of the number of species within a given region. This analysis of 35 cetaceans shows the probability of occurrence based on relative environmental suitability, including species preferences for water temperature, depth, salinity, and distance to land along the California Bight. 


```{r, echo = TRUE, results = "hide"}
coastline <- ne_download(scale = 50, type = 'land', category = 'physical', returnclass = "sf")
```


```{r}
cetaceans <- list.files(path = here("ca_cetaceans"), full.names = TRUE)


#create a stack of all rasters
cetaceans_rast <- raster::stack(cetaceans)

# create the vector containing elements to make a matrix
reclass <- c(0, 0.59, 0, 0.6, 1, 1)

#make the matrix out of the vector
matrix_reclass <- matrix(reclass, ncol = 3, byrow = TRUE)

#reclassify the rasters in the stack based on the threshold
presence_rast <- reclassify(cetaceans_rast, matrix_reclass)

# add up all the values from the reclassified rasters into a single raster
rast_sum <- raster::calc(cetaceans_rast, fun = sum, na.rm = TRUE)

# create data frame
species_sum_df <-  raster::rasterToPoints(rast_sum) %>% 
  as.data.frame()


```

```{r, fig.cap= "Cetaceans species richness based on species occurence probability data from AquaMaps. Species presence is determined by a probability threshold of 0.6 or greater for individual species, and summed over all 35 species to create a raster of richness."}

ggplot()+
  geom_raster(data = species_sum_df, aes(x = x, y = y, fill = layer))+
  scale_fill_gradient("Number of Species", low = "darkblue", high = "cyan2") +
  geom_sf(data = coastline, fill = "burlywood1")+
  coord_sf(xlim = c(-124.8,-115.5), ylim = c(32,38)) +
  theme_minimal()+
  labs(
    x = " ",
    y = " "
  )
  
```


### Data Citation

Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org

Made with Natural Earth.

