---
title: "Oil Spill Spatial Analysis"
description: |
  Spatial analysis of oil spills in California.
author:
  - name: Anastasia Kunz
date: 02-21-2021
output:
  distill::distill_article:
    self_contained: false
    theme: flatly
    code_folding: hide
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(tmap)
library(here)
library(sf)
library(janitor)
library(lubridate)
library(viridis)
```

## Oil Spills Spatial Exploration

Explore occurrence data for California oil spills in 2008 through the interactive map below.

```{r}
# Read in the data:
oilspill <- read_sf(here("oil_spill_data"),
                    layer = "Oil_Spill_Incident_Tracking_%5Bds394%5D") %>% 
  clean_names() %>% 
  mutate(date = as.Date(dateofinci)) %>% 
  mutate(date = ymd(date)) %>% 
  rename(name = localecoun) %>% 
  select(name, date, inlandmari)

counties <- read_sf(here("ca_counties"),
                    layer = "CA_Counties_TIGER2016") %>% 
  clean_names() %>% 
  select(name)


```

```{r, include=FALSE}

# Check projection
st_crs(oilspill)

# Set Counties CRS
counties <- st_transform(counties, st_crs(oilspill))



```


```{r}

tmap_mode("view")

tm_shape(counties)+
  tm_polygons(alpha = 0.4)+
  tm_shape(oilspill)+
  tm_dots(col = "red", scale = .8, alpha = 0.6)+
  tm_basemap("Esri.WorldTopoMap")
  

```




```{r}
oil_counties <- counties %>% 
  st_join(oilspill) %>% 
  count(name.x, inlandmari) %>% 
  filter(inlandmari == "Inland")
```


## County Oil Spills Choropleth Map

Some California counties have much higher concentrations of inland oil spills than others. The choropleth map below shows the number of inland oil spills for each California county. 

```{r}

ggplot()+
  
  geom_sf(data = counties)+
  geom_sf(data = oil_counties, aes(fill = n))+
  scale_fill_viridis("Number of Oilspills", option = "plasma")+
  theme_void()+
  labs(
    title = "Inland Oilspills in California by County"
  )
  



```



### Data Citation

California Department of Fish and Game, Office of Spill Prevention and Response. Oil Spill Incident Tracking [ds394] (2008).



